Hydra Battles and Cie

### 2021-07-27 doc/chapter-primrec.tex
start to add snippets to chapter 2 (hydras)

In particular, the library \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
MoreOver, the library \href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems} contains 
a proof that the length of an hydra battle (according to the initial replication factor) is not primitive recursive in general.

@@

The formal definition of primitive recursive functions lies in the library
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec},
with preliminary definitions in 
\href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}
and
\href{../theories/html/hydras.Ackermann.misc.html}{Ackermann.misc}.

@@

\textbf{Hint:} You may have to look again at the lemmas of the library
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} if you meet some difficulty.
You may start this exercise with the file
    \href{https://https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MorePRExamples.v}{exercises/primrec/MorePRExamples.v}.
\end{exercise}

@@

\emph{You may start this exercise with
    \href{https://https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MinPR.v}{exercises/primrec/MinPR.v}.}

>>>>

In particular, the library \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/MoreAck/AckNotPR.v}{MoreAck.AckNotPR} contains the well-known  proof that the Ackermann function is not primitive recursive (see Section~\vref{sect:ack-not-PR}).
Moreover, the library \href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Hydra/Hydra_Theorems.v}{Hydra.Hydra\_Theorems} contains 
a proof that the length of an hydra battle (according to the initial replication factor) is not primitive recursive in general.

@@

The formal definition of primitive recursive functions lies in the library
\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Ackermann/primRec.v}{Ackermann.primRec},
with preliminary definitions in 
\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Ackermann/extEqualNat.v}{Ackermann.extEqualNat}
and
\href{https://github.com/coq-community/hydra-battles/blob/master/theories/ordinals/Ackermann/misc.v}{Ackermann.misc}.

@@

\textbf{Hint:} You may have to look again at the lemmas of the library
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec} if you meet some difficulty.
You may start this exercise with the file
    \href{https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MorePRExamples.v}{exercises/primrec/MorePRExamples.v}.
\end{exercise}

@@

\emph{You may start this exercise with
    \href{https://github.com/coq-community/hydra-battles/blob/master/exercises/primrec/MinPR.v}{exercises/primrec/MinPR.v}.}

### 2021-07-27 doc/part-hydras.tex
start to add snippets to chapter 2 (hydras)

\begin{Coqsrc}
Inductive Hydra : Set :=
| node :  Hydrae -> Hydra
with Hydrae : Set :=
| hnil : Hydrae
| hcons : Hydra -> Hydrae -> Hydrae.
\end{Coqsrc}

@@

\begin{Coqalt}
Module Alt.

Inductive Hydra: Set :=
  hnode (daughters : list Hydra).

End Alt.
\end{Coqalt}

@@

\begin{Coqsrc}
(** heads *)
Notation head := (node hnil).

(** nodes  with 1, 2 or 3 daughters *)
Notation hyd1 h := (node (hcons h hnil)).
Notation hyd2 h h' := (node (hcons h (hcons h' hnil))).
Notation hyd3 h h' h'' := 
                   (node (hcons h (hcons h' (hcons h'' hnil)))).
\end{Coqsrc}

@@

\begin{Coqsrc}
Example Hy := hyd3 head
                   (hyd2
                      (hyd1 
                         (hyd2 head head))
                      head) 
                   head.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint hcons_mult (h:Hydra)(n:nat)(s:Hydrae):Hydrae :=
  match n with 
  | O => s
  | S p => hcons h (hcons_mult h p s)
  end.

(** A node with n clones of the same daughter *)

Definition hyd_mult h n :=
  node (hcons_mult h n hnil).
\end{Coqsrc}

@@

\begin{Coqsrc}
Example Hy'' := 
     hyd2 head
          (hyd2 (hyd_mult (hyd1 head) 5)
                head).
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint hsize (h:Hydra) : nat :=
  match h with node l => S (lhsize l)
  end
with lhsize l : nat :=
  match l with hnil => 0
            | hcons h hs => hsize h + lhsize hs 
  end.

 Compute hsize Hy.
\end{Coqsrc}

\begin{Coqanswer}
 = 9
     : nat 
\end{Coqanswer}

@@

\begin{Coqsrc}
Fixpoint height  (h:Hydra) : nat :=
  match h with node l => lheight l
  end
with lheight l : nat :=
  match l with 
  | hnil => 0
  | hcons h hs => Max.max (S (height h)) (lheight hs)
  end.
\end{Coqsrc}

\begin{Coqsrc}
Compute height Hy.
\end{Coqsrc}

\begin{Coqanswer}
 = 4
     : nat  
\end{Coqanswer}

@@

\begin{Coqanswer}
Hydra_ind :
forall P : Hydra -> Prop,
(forall h : Hydrae, P (node h)) -> forall h : Hydra, P h
\end{Coqanswer}

Ler us start a simple proof by induction.

@@

Lemma height_lt_size (h:Hydra) :
  height h <= hsize h.
Proof.
  induction h as [s].
\end{Coqbad}

\begin{Coqanswer}
1 subgoal, subgoal 1 (ID 11)

  s : Hydrae
  ============================
   height (node s) <= hsize (node s)
\end{Coqanswer}

@@

% \begin{Coqbad}
% induction  s as [ | h s'].
%   -   cbn; auto with arith.
%   - 
% \end{Coqbad}

\begin{Coqanswer}
 1 focused subgoal
(unfocused: 0), subgoal 1 (ID 19)

  h : Hydra
  s' : Hydrae
  IHs' : height (node s') <= hsize (node s')
  ============================
   height (node (hcons h s')) <= hsize (node (hcons h s'))

Note that the displayed subgoal does not contain any assumption on \texttt{h}, thus there is no way to 
infer any property about the height and size of the hydra \texttt{(hcons h t)}.

\begin{Coqbad}
Abort.

End Bad.
\end{Coqbad}

>>>>

\input{movies/snippets/Hydra_Definitions/HydraDef}

@@

\input{movies/snippets/Hydra_Definitions/HydraAlt}

@@

\input{movies/snippets/Hydra_Definitions/headsEtc}

@@

\input{movies/snippets/Hydra_Examples/Hy}

@@

\input{movies/snippets/Hydra_Definitions/hconsMult}

@@

\input{movies/snippets/Hydra_Examples/HySecond}

@@

\input{movies/snippets/Hydra_Definitions/hsize}

\input{movies/snippets/Hydra_Examples/HySize}

@@

\input{movies/snippets/Hydra_Definitions/height}

\input{movies/snippets/Hydra_Examples/HyHeight}

@@

\input{movies/snippets/Hydra_Examples/HydraInd}

Let us start a simple proof by induction.

@@

\input{movies/snippets/Hydra_Examples/BadInductiona}

@@

\input{movies/snippets/Hydra_Examples/BadInductionb}

The first subgoal is trivial.

\input{movies/snippets/Hydra_Examples/BadInductionc}

Let us look at the second sub-goal of the induction.

\input{movies/snippets/Hydra_Examples/BadInductiond}

We notice that this subgoal does not contain any hypothesis
on the height and size of the hydra \texttt{h}. So, it looks hard to prove the conclusion. Let's stop.

\input{movies/snippets/Hydra_Examples/BadInductione}
